Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-04-23 manuscript/manuscript.tex
updated ms
update ms

In analyzing both the empirical datasets included in our study and our
simulation results, we noticed the temporal covariances and $G(t)$ statistics
are sensitive to how fixations (i.e. when a site is polymorphic but is fixed or
lost during the course of its trajectory) are handled. Through simulations and
comparing the results of analyzing our empirical datasets both ways, we have
found trade-offs to both approaches, but overall find that not removing fixed
sites is a more conservative, safer approach, and is what we use throughout our
paper.

@@

Still, we wanted to ensure that our empirical re-analyses were robust to our
choice of how we handled fixed sites. Given that under some simulation
conditions, we find that spurious negative covariances can occur when fixations
are included, we wanted to ensure that our result that the temporal covariances
of \textcite{Barghi2019-qy} that become negative at later timepoints are not an
artifact of the way fixations are handled. First, we regenerated Figure 1 A and
B, except with temporal covariances and $G(t)$ calculated excluding fixed or
lost sites. We find the same qualitative results regardless of whether we
include or exclude fixed/lost sites; importantly, we still see a significant
signal that the temporal covariances at later timepoints become negative.

@@

  \caption{A version of Figure 1 A and B where sample frequencies of zero or
    one are ignored, e.g. due to fixation, are ignored. Like the original
    Figure 1 in the main text, the 95\% pivot confidence intervals are via a
    bootstrapping procedure; note that we find excluding fixed sites leads to
  more asymmetric confidence intervals. Overall, we find that despite handling
fixed/lost sites differently, we still find significant negative temporal
covariances at later timepoints.}

>>>>

In analyzing both the empirical datasets included in our study and our
simulation results, we noticed the temporal covariances and $G(t)$ statistics
can differ depending on how allele frequencies of zero or one are handled.
Generally, temporal covariances should be calculated on polymorphic sites; once
a site has reached fixation or loss, its allele frequency change $\Delta p_t =
0$ and including these sites in the temporal covariance calculation can lead to
biases, which we discuss below. However, with \emph{sample} allele frequencies,
rather than population frequencies, a site with observed frequency zero or one
may still be segregating, but by chance not sampled at a timepoint. Here, we
discuss the effect of including sites with frequency zero or one, and show our
empirical results are not qualitatively different when analyzed excluding fixed
sites.

First, we found that under strong truncation selection (i.e. the number of
selected individuals is small), our measure $G(t)$ can become negative. This
possibility was discussed in \textcite{Buffalo2019-io} as a result of
fluctuating selection; here, however, the buildup of negative covariances
driving this

Since our simulations were focused on understanding how genetic architecture
and mode of selection impact temporal covariance and $G(t)$ (with a limited
number of neutral sites, due to computational limitations), population
frequencies were used.

If frequencies are measured without sampling noise, including fixed or lost
sites in the calculation of temporal covariance and $G(t)$ has the effect of
shrinking the covariance estimates towards zero,

how fixations (i.e. when a site is polymorphic but is fixed or
lost during the course of its trajectory) are handled. Through simulations and
comparing the results of analyzing our empirical datasets both ways, we have
found trade-offs to both approaches, but overall find that not removing fixed
sites is a more conservative, safer approach, and is what we use throughout our
paper.

@@

Still, we wanted to ensure that our empirical re-analyses were robust to our
choice of how we handled fixed sites. Given that under some simulation
conditions, we find that spurious negative covariances can occur when fixations
are included, we wanted to ensure that our result that the temporal covariances
of \textcite{Barghi2019-qy} that become negative at later timepoints are not an
artifact of the way fixations are handled. First, we regenerated Figure 1 A and
B, except with temporal covariances and $G(t)$ calculated excluding fixed or
lost sites. We find the same qualitative results regardless of whether we
include or exclude fixed/lost sites; importantly, we still see a significant
signal that the temporal covariances at later timepoints become negative. As
expected the temporal covariances are lessened by excluding fixed sites.

@@

  \caption{The effect of including fixed/lost sites in the calculation of the
  temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
data. Dashed lines are those including fixed/lost sites (i.e. the original
Figure 1), and solid lines are excluding fixed/lost sites.}
